A community-maintained standard library of population genetic models

### 2019-12-19 popSimManu.tex
added comments and minor edits

\documentclass[12pt,halfline,a4paper]{ouparticle}
\usepackage{amsmath,amssymb}
\usepackage[round]{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage{siunitx}

@@

\newcommand{\Stdpopsim}{\texttt{Stdpopsim}\xspace}
\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairwayplot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal2}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

@@

\title{A community-maintained standard library of population genetic simulations}
% First authors
\author[1,$\star$]{Jeffrey R. Adrion}
\author[2,$\star$]{Christopher B. Cole}
\author[3,$\star$]{Noah Dukler}
\author[1,$\star$]{Jared G. Galloway}
\author[4,$\star$]{Ariella L. Gladstein}
\author[5,$\star$]{Graham Gower}
\author[6,$\star$]{Christopher C. Kyriazis}
\author[7,$\star$]{Aaron P. Ragsdale}
\author[8,$\star$]{Georgia Tsambos}
% Second Authors
\author[9]{Franz Baumdicker}
\author[10]{Jedidiah Carlson}
\author[11]{Reed A. Cartwright}
\author[12]{Arun Durvasula}
\author[13]{Bernard Y. Kim}
\author[14]{Patrick McKenzie}
\author[15]{Philipp W. Messer}
\author[16]{Ekaterina Noskova}
\author[17]{Diego Ortega-Del Vecchyo}
\author[5]{Fernando Racimo}
\author[18]{Travis J. Struck}
%Senior Authors
\author[7,$\dagger$]{Simon Gravel}
\author[18,$\dagger$]{Ryan N. Gutenkunst}
\author[6,$\dagger$]{Kirk E. Lohmeuller}
\author[1,$\dagger$]{Peter L. Ralph}
\author[4,$\dagger$]{Daniel R. Schrider}
\author[3,$\dagger$]{Adam Siepel}
\author[19,$\dagger$,$\aleph$]{Jerome Kelleher}
\author[1,$\dagger$,$\aleph$]{Andrew D. Kern}

@@

Assessing accuracy in population genetics is challenging in large part because
 ``ground-truth'' in population genetics comes from simulation,
not from empirical observations, and as a field there has been no agreement
on community standards or best practices. The general modus operandi to date has been to
introduce methods that are validated by a set of bespoke simulations produced by individual
groups. Beyond inefficiency from duplicated effort, such a situation leads
to decreased reproducibility and transparency across the entire field.
Here we describe \stdpopsim, a community-driven catalog
of standardized population genetics models spanning a variety of common study organisms.
This resource will empower the field to rigorously evaluate population genetic software,
%and select the best tool for a given task,
thereby substantially increasing the reliability of population genetic inferences.
%DRS: still not in love with this last sentence here, but tried to make it a little less sappy by adding some explanation.
%ADK: made some edits here
%RNG: Removed the sap. :-P Feel free to put it back if you like it better.
%JK: not crazy about it either. Making it shorter helps though?

It has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect the ultimate results of population
genetic inference. Yet there
have been several clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, it is reasonable
to assume generally that specific methods will perform better than others under certain scenarios.
However, because these conditions are typically unknown, empirical researchers
lack principled guidelines for deciding which statistical method is best suited
to accurately answer their particular question. The need for empirical
guidance will only increase as researchers continue to generate more genomic
variation data from non-model taxa and wish to make inferences regarding
demography and selection.

One way to benchmark different methods for statistical inference in population
genetics is to apply the methods to simulated datasets where the true model
parameters are known. Methods that perform well will accurately infer these
parameters and will also provide reasonable measures of uncertainty of the
estimated parameters. Indeed, simulation studies are a standard ingredient in
most papers proposing new methods in population genetics. However, these
studies often use simulations of the particular models for which the method was designed,
which showcases the novel method rather than rigorously comparing performance to competing methods.
Of course, testing a new method in the situation for which it was designed is an important step,
but more rigorous testing is necessary for tools that are to be widely used in the field.
% Moreover researchers are generally forced to reimplement previously
% inferred scenarios independently, which makes good benchmarking
% a major additional time investment, and can lead to errors.
% Lastly, the degree to which different methods are compared can be quite variable.
Many more researchers in the broader field
use simulations to evaluate statistical power and perform other exploratory analyses.
The many steps of implementing a moderately complicated population genetics simulation,
including translating published demographic models into input for a simulator
and obtaining genetic maps,
present a major obstacle, especially to new researchers.

% One reason why simulation studies do not compare the performance of many methods
% across a wide range of situations
% is that simulation studies are challenging to carry out, for several reasons.
% First, while there are many published population genetic models available, their
% parameters are often not easily accessible, usually presented in supplemental
% tables. Translating these parameters into input for a given simulation engine can
% be a time-consuming and error prone process. In addition, papers often present many
% different models, and different authors may choose to select distinct models,
% making comparison difficult. While improvements in simulation software and ever
% faster computers have partially ameliorated this concern, generating the
% simulated datasets can still be a time and computing resource bottleneck.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains five organisms: humans, \emph{Pongo~pygmaeus},
\emph{Drosophila~melanogaster}, \emph{Arabidopsis~thaliana}, and \emph{Escherichia~coli}. For each
% CK: i think we should spell out genus at first mention.
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.

The software has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/Figure1.pdf}
\caption{\textbf{Structure of \stdpopsim}. \textbf{A.} The
hierarchical organization of the \stdpopsim catalog contains all model simulation information
within individual species (expanded information shown here for \emph{Homo sapiens} only). 
Each species is associated with a representation of the physical genome, and one or more genetic maps and demographic models.
Dotted lines indicate that only a subset of these categories is shown. \textbf{B.} Example code to specify
and simulate models using \stdpopsim via the API (top) or\textbf{C.} through a command line interface.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The first contribution of the PopSim consortium is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{P.~pygmaeus},
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as genetic maps, which model observed
heterogeneity in recombination rate along chromosomes. As such maps are often large,
we do not distribute them directly with the software, but make them available
for download in a standard format. When a simulation using such a map is
requested by the user, \stdpopsim will transparently download the map
data into a local cache (\texttt{$\sim$/.local/cache/stdpopsim} by default
on Unix platforms), where it can be quickly retrieved for subsequent
simulations.
In the initial version of \stdpopsim we support
the HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps in humans;
the \cite{salome2011recombination} map in \emph{A.~thaliana};
and the \cite{comeron2012many} map in \emph{D.~melanogaster}.
The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic variation. (See the Methods for a description of the community
development and quality-control process for these models.)

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \Stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim} as
an additional backend.

@@

\subsection*{The \stdpopsim Catalog}
The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Pongo}, \emph{Drosophila}, and \emph{Arabidopsis}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

At the time of writing, more models of human history have been implemented than for
any other organism. These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth), the three population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations, the \cite{tennessen2012evolution} two population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth, the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components, a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture, a complex model of ancient eurasian admixture from \cite{kamm2019efficiently},
and a synthetic model of oscillating population size from \cite{schiffels2014inferring}.
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{D. melanogaster}, three from \emph{A. thaliana}, an one each from \emph{P. pygmaeus} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample, as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations. For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes. We have also implemented a two epoch and a three epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene}. For \emph{P. pygmaeus} we have implemented the two
species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
and \emph{P. abelii}. For \emph{E. coli} we have a simple, large constant population size model implemented.
Beyond organism specific models \stdpopsim also includes a generic piecewise constant size model and
isolation with migration (IM) model which can be used with any genome and genetic map combination.

\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

@@

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above. The pipeline was generalized to
run $R$ replicates with $C$ chromosomes for $n$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the pipeline prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairwayplot}, and
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the \cite{ragsdale2019models} model and using the
HapMapII genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under under a sophisticated demographic model
and empirical genetic map and a model of human migration out of Africa
that includes archaic admiture \citep{ragsdale2019models}. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC).
There is no single ``true'' reference for effective population size
because of model misspecification -- the inference methods are fitting a single population model
to data simulated from mulitple populations.
However, many methods work by matching coalescence time distributions,
and a single-population model with varying population size can match any coalescence time distribution
(in which case coalescence rate is the inverse of the effective size).
For this reason, solid black lines do not show historical census sizes,
but rather inverse coalescence rates calculated analytically in \texttt{msprime} (see Appendix).
Blue lines show estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
generally the methods are accurate for this model of human history.

\Stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but a different demographic
history that does not include archaic admixture \citep{gutenkunst2009inferring}. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

To see how well methods might do at recovering the population history of a constant-sized population,
we can simulate under a human genome architecture and genetic map, but without demography.
We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
are recovering a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/d_mel_Sheehan_mask2.png}
\caption{\textbf{Comparing estimates of $N(t)$ in \emph{Drosophila}}. Population
size over time ($N(t)$) estimated from an African population sample. Data were generated by simulating
replicate \emph{D.melanogaster} genomes under the three-epoch \cite{sheehan2016deep} model
with the genetic map inferred in \cite{comeron2012many}. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_sheehan}
\end{center}
\end{figure}

As most method development for population genetics has been focused on human
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from a demographic
model estimated from an African sample of \emph{D. melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting and generally worse than what we
observe under simulations of the human genome.

>>>>

\documentclass[12pt,halfline,a4paper]{ouparticle}
\usepackage{amsmath,amssymb}
\usepackage[round]{natbib}
\usepackage{graphicx,rotating}
\usepackage{dsfont}
\usepackage{xspace}
\usepackage{booktabs}
\usepackage[binary-units]{siunitx}
\usepackage[section]{placeins}

@@

\newcommand{\Stdpopsim}{\texttt{Stdpopsim}\xspace}
\newcommand{\stdpopsim}{\texttt{stdpopsim}\xspace}
\newcommand{\dadi}{$\partial a \partial i$\xspace}
\newcommand{\MSMC}{\texttt{MSMC}\xspace}
\newcommand{\smcpp}{\texttt{smc++}\xspace}
\newcommand{\stairwayplot}{\texttt{stairway plot}\xspace}
\newcommand{\fastsimcoal}{\texttt{fastsimcoal2}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

@@

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\rhead{\texttt{stdpopsim}}
\lhead{PopSim Consortium}
\fancyfoot[C]{\thepage}

@@

\title{A community-maintained standard library of population genetic models}
% First authors
\author[1,$\star$]{Jeffrey R. Adrion}
\author[2,$\star$]{Christopher B. Cole}
\author[3,$\star$]{Noah Dukler}
\author[1,$\star$]{Jared G. Galloway}
\author[4,$\star$]{Ariella L. Gladstein}
\author[5,$\star$]{Graham Gower}
\author[6,$\star$]{Christopher C. Kyriazis}
\author[7,$\star$]{Aaron P. Ragsdale}
\author[8,$\star$]{Georgia Tsambos}
% Second Authors
\author[9]{Franz Baumdicker}
\author[10]{Jedidiah Carlson}
\author[11]{Reed A. Cartwright}
\author[12]{Arun Durvasula}
\author[13]{Bernard Y. Kim}
\author[14]{Patrick McKenzie}
\author[15]{Philipp W. Messer}
\author[16]{Ekaterina Noskova}
\author[17]{Diego Ortega-Del Vecchyo}
\author[5]{Fernando Racimo}
\author[18]{Travis J. Struck}
%Senior Authors
\author[7,$\dagger$]{Simon Gravel}
\author[18,$\dagger$]{Ryan N. Gutenkunst}
\author[6,$\dagger$]{Kirk E. Lohmeuller}
\author[1,$\dagger$]{Peter L. Ralph}
\author[4,$\dagger$]{Daniel R. Schrider}
\author[3,$\dagger$]{Adam Siepel}
\author[19,$\dagger$,$\aleph$]{Jerome Kelleher}
\author[1,$\dagger$,$\aleph$]{Andrew D. Kern}

@@

Assessing the accuracy of inference methods for population genetics is
challenging in large part because the ``ground-truth'' in question
generally comes not from direct empirical observations, as the relevant
historical processes can rarely be observed, but instead from simulations.
Population genetic simulations are therefore critically important to the
field, yet there has been no systematic attempt to establish community
standards or best practices for executing them. Instead, the general modus
operandi to date has been for individual groups to validate their own
methods using bespoke simulations. 
Often these studies focus more on showcasing a novel method than on
Moreover, this approach results in a great deal of duplicated effort across research
It is also a barrier to entry to the field, because new researchers can
struggle with 
the many steps involved
in implementing a state-of-the-art population genetics simulation,
including identifying appropriate demographic models from the literature, 
translating them into input for a simulator, obtaining appropriate genetic
maps, and choosing appropriate values for key population genetic parameters.
groups, and contributes to decreased reproducibility and transparency.
rigorously comparing it with competing methods.

A related issue is that it has been challenging to assess the degree to which modeling assumptions
and choices of data summaries can affect population
genetic inferences. Yet there
are clear examples of different methods yielding fundamentally
different conclusions. For example, Markovian coalescent methods applied to human genomes have
suggested large ancient ($>100,000$ years ago) ancestral population sizes and
bottlenecks that have not been detected by methods based on allele frequency spectra
\citep[see][]{beichman2017comparison}.
These distinct methods differ in how they model, summarize, and optimize fit to
genetic variation data, suggesting that such design choices can greatly affect the
performance of the inference. Furthermore, some methods are likely to
perform better than others under certain scenarios, but 
researchers lack principled guidelines for selecting the best method for addressing 
their particular questions. The need for empirical
guidance will only increase as researchers 
seek to apply population genetic methods to a growing collection of non-model taxa.

For these reasons, we have generated a standardized, community-driven resource
for simulating published demographic models from a number of popular study systems.
This resource, which we call \stdpopsim, makes running
realistic simulations for population genetic analysis a simple matter of
choosing pre-implemented models from a community-maintained catalog.
The \stdpopsim catalog currently contains three organisms: humans,
\emph{Drosophila~melanogaster}, and \emph{Arabidopsis~thaliana}. For each
organism the catalog contains details on the physical organization (e.g., chromosome structure)
of their genomes, one or more genetic maps, default population-level parameters (mutation rate,
generation time) and one or more published demographic histories. Through
either a command line interface or a simple Python API, users can specify which
organism, genetic map, chromosome, and demographic history they are interested in simulating, and the
simulation output from their chosen model is returned.
In this way, \stdpopsim will enable rigorous evaluation of population genetic software,
and contribute to increased reliability of population genetic inferences.

The \stdpopsim library has been developed by the PopSim Consortium using a
distributed open source model, with strong procedures in place
to continue its growth and maintain its quality.
Importantly, we have rigorous quality control methods to ensure implemented models are accurate
and have documented methods for others to contribute new modules.
We invite new collaborators to join our community.
Below we describe the resource and give
examples of how it can be used to benchmark demographic inference methods.

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\linewidth]{display_items/Figure1.pdf}
    \caption{\textbf{Structure of \stdpopsim}. \textbf{(A)} The
hierarchical organization of the \stdpopsim catalog contains all model simulation information
within individual species (expanded information shown here for \emph{H. sapiens} only). 
Each species is associated with a representation of the physical genome, and one or more genetic maps and demographic models.
Dotted lines indicate that only a subset of these categories is shown. \textbf{(B)} Example code to specify
and simulate models using \stdpopsim via the API or \textbf{(C)} through a command line interface.}
\label{fig:cartoon}
\end{center}
\end{figure}

\section*{Results}
\paragraph{The \stdpopsim library}
The first contribution of the PopSim consortium is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans,
\emph{D.~melanogaster}, \emph{A.~thaliana}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as genetic maps, which model observed
heterogeneity in recombination rate along chromosomes. As such maps are often large,
we do not distribute them directly with the software, but make them available
for download in a standard format. When a simulation using such a map is
requested by the user, \stdpopsim will transparently download the map
data into a local cache (\texttt{$\sim$/.local/cache/stdpopsim} by default
on Unix platforms), where it can be quickly retrieved for subsequent
simulations.
In the initial version of \stdpopsim we support
the HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps in humans;
the \cite{salome2011recombination} map in \emph{A.~thaliana};
and the \cite{comeron2012many} map in \emph{D.~melanogaster}.
The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic variation. (See the Methods for a description of the community
development and quality-control process for these models.)

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. 
We have implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim}
as an alternative backend,
to allow simulation of processes that cannot be modeled under the coalescent.

@@

\subsection*{The Species Catalog}
The central feature of \stdpopsim is the species catalog, a systematic organization
of the key quantitative data needed to simulate a given species. These include a description
of the assembly, information about mutation rate, recombination rate(s), and generation time
in addition to a series of demographic models that are specific to that organism.

The current contents of the \stdpopsim catalog are shown in \autoref{tab:catalog}. At the time of
writing, the PopSim Consortium has implemented previously published demographic models for
humans, \emph{Drosophila}, and \emph{Arabidopsis}, 
and we have models in development for \emph{Pongo abelii} and \emph{Escherichia coli}.
These range from
simple, single population histories \cite[e.g.,][]{sheehan2016deep},
to complex models which include population splitting, migration, and archaic
admixture \cite[e.g.,][]{ragsdale2019models}.

@@

Currently, more models of human history have been implemented than for
any other organism (see Table \ref{tab:catalog}). 
These models include: a simplified version of the \cite{tennessen2012evolution}
model with only the African population specified (expansion from the ancestral
population and recent growth; Africa\_1T12), the three-population model of \cite{gutenkunst2009inferring}
which specifies the out-of-Africa bottleneck as well as the subsequent divergence of
the European and Asian populations (OutOfAfrica\_3G09), the \cite{tennessen2012evolution} two-population variant of the
Gutenkunst model which does not include Asian populations, but more explicitly models
recent rapid human population growth (OutOfAfrica\_2T12), the \cite{browning2018ancestry} admixture model
for American populations which specifies ancestral African, European, and Asian population
components (AmericanAdmixture\_4B11), a three-population out-of-Africa model from \cite{ragsdale2019models}
which includes archaic admixture (OutOfAfricaArchaicAdmixture\_5R19),
a complex model of ancient Eurasian admixture from \cite{kamm2019efficiently} (AncientEurasia\_9K19),
and a synthetic model of oscillating population size from \cite{schiffels2014inferring} (Zigzag\_1S14).
Together these models
contain features believed to have widespread impacts in real data (e.g. bottlenecks, population growth,
admixture) and are therefore highly pertinent in the context of method development.

For non-human genomes we have have implemented two demographic histories for
\emph{D. melanogaster}, three from \emph{A. thaliana}, and have under development
model for \emph{P. abelii} and \emph{E. coli}.
For \emph{D. melanogaster} we have implemented the three-epoch model estimated by \cite{sheehan2016deep} from
an African sample (African3Epoch\_1S16), as well as the out-of-Africa divergence
and associated bottleneck model of \cite{li2006inferring} which jointly models African
and European populations (OutOfAfrica\_2L06). For \emph{A. thaliana}, we implemented the
model in \cite{durvasula2017african} inferred using \MSMC. This model includes
a continuous change in population size over time, rather than pre-specified epochs of different
population sizes (SouthMiddleAtlas\_1D17). We have also implemented a two-epoch and a three-epoch model estimated from African
samples of \emph{A. thaliana} in \cite{huber2018gene} (African2Epoch\_1H18 and African3Epoch\_1H18). 
%For \emph{P. pygmaeus} we have implemented the two
%species divergence model of \cite{locke2011comparative}, which models isolation with migration between \emph{P. pygmaeus}
%and \emph{P. abelii}. 
%For \emph{E. coli} we have a simple, large constant population size model implemented.
Beyond organism specific models \stdpopsim also includes a generic piecewise constant size model and
isolation with migration (IM) model which can be used with any genome and genetic map combination.

We note that we have put a good deal of effort into coming up with a standardized naming
conventions for species, genetic maps, and demographic models that will enable long term stability
of unique identifiers used throughout \stdpopsim. This is an essential feature to guarantee reproducibility.
\subsection*{Use case: comparing methods of demographic inference}
As an example of the utility of \stdpopsim, we demonstrate how it can be used
to simply and fairly compare popular demographic inference methods.
Although we present comparison of results from several
methods, our aim at this stage is not to provide an exhaustive
evaluation or ranking of these methods. Our hope is instead that future work built upon this resource
will enable more detailed exploration of the strengths and weaknesses of the numerous
inference methods that are available to the population genetics community.

@@

For single-population population size histories, we compared
\MSMC~\citep{schiffels2014inferring}, \smcpp~\citep{terhorst2017robust}, and
\stairwayplot~\citep{liu2015exploring}
 on simulated genomes sampled from a single population,
in a number of the demographic models described above. The workflow was generalized to
run $R$ replicates with $C$ chromosomes for $n$ samples, for a total of $R \times C$
simulations for each demographic model. After simulation, the workfow prepares
input files for each of the respective inference methods by grouping all
chromosomes, for each sample, into a single file. For each of the $R$ simulation replicates, this step results in an
input file for each of the
respective inference methods and derived from the same simulated tree sequences.
Each of the inference programs are then run in parallel, followed by plotting of
$N(t)$ estimates from each program.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/homo_sapiens_mask_Ragsdale.png}
\caption{\textbf{Comparing estimates of $N(t)$ in humans}. Here we show estimates of population
size over time ($N(t)$) inferred using 4 different methods: \smcpp, \texttt{stairway plot}, and
\MSMC with $n=2$ and $n=8$. Data were generated by simulating
replicate human genomes under the OutOfAfricaArchaicAdmixture\_5R19 model and using the
HapMapII\_GRCh37 genetic map~\citep{international2007second}. From top to bottom we show estimates for each
of the three populations in the model (YRI, CEU, and CHB). In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_ragsdale}
\end{center}
\end{figure}

% JK here --- I think it would really help if we described what the models
% were, briefly, rather than saying who inferred them.
In Figure \ref{fig:n_t_ragsdale} we present results from our $N(t)$ analysis pipeline
run on whole genome simulations under under OutOfAfricaArchaicAdmixture\_5R19,
a model of human migration out of Africa that includes archaic admixture 
\citep{ragsdale2019models}, along with an empirical genetic map. In each column of this figure
we show inferred $N(t)$ for the three extant populations in the model.
In each row we show comparisons among the methods (note that we compare two differing
sample sizes for \MSMC).
% ALG: I still think this sentence is confusing.
There is no single ``true'' reference for effective population size
because of model misspecification---the inference methods are fitting a single population model
to data simulated from mulitple populations.
However, many methods work by matching coalescence time distributions,
and a single-population model with varying population size can match any coalescence time distribution
(in which case coalescence rate is the inverse of the effective size).
For this reason, solid black lines do not show historical census sizes,
but rather inverse coalescence rates calculated analytically in \texttt{msprime} (see Appendix).
Blue lines show estimates from each of three replicate whole genome simulations.
While there is variation in accuracy among methods, populations, and individual replicates,
generally the methods are accurate for this model of human history.

\stdpopsim allows us to readily compare these estimates to those drawn from a different
model of human history. In Figure \ref{fig:n_t_gutenkunst} we show estimates of
$N(t)$ from simulations using the same physical and genetic maps, but from the OutOfAfrica\_3G09
demographic model that does not include archaic admixture. Again we see that each
of the methods is capturing relevant parts of the population history, although the
discrepancy between inferred and simulated values varies with $t$. In comparing inferences between the
models it is interesting to note that $N(t)$ estimates for the CHB and CEU
simulated populations are generally better across methods than estimates from the YRI
simulated population.

To see how well methods might do at recovering the population history of a constant-sized population,
we can simulate under a human genome architecture and genetic map, but without demography.
We show results of such an
experiment in Figure \ref{fig:n_t_HomSap_constant}. Again we see that all methods
recover a modern population size within a factor of two of the truth, however
SMC-based methods, perhaps due to their regularization, tend to learn sinusoidal
patterns of population size even though no change is present.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{display_items/d_mel_Sheehan_mask2.png}
\caption{\textbf{Comparing estimates of $N(t)$ in \emph{Drosophila}}. Population
size over time ($N(t)$) estimated from an African population sample. Data were generated by simulating
replicate \emph{D.melanogaster} genomes under the African3Epoch\_1S16 model
with the genetic map inferred in \cite{comeron2012many}. In shades of blue we show the estimated
$N(t)$ trajectories for each replicate. In black we show the true population size history as inferred
for the rate of coalescence in the demographic model.}
\label{fig:n_t_sheehan}
\end{center}
\end{figure}

As most method development for population genetics has been focused on human
data, it is of consequence to ask how such methods might perform in non-human
genomes. Figure \ref{fig:n_t_sheehan} shows parameter estimates from the African3Epoch\_1S16
model, originally estimated from an African sample of \emph{D. melanogaster} \citep{sheehan2016deep}.
As with humans, we use \stdpopsim to simulate replicate \emph{Drosophila} genomes under
an empirically derived genetic map, and then try to infer back the parameters of the model.
Accuracy is mixed among methods in this setting and generally worse than what we
observe under simulations of the human genome.
